Multi-fluids for Representing Sub-filter-scale Convection

### 2020-03-23 submit1/manuscript.tex
First submmitted version and empty revised version and responses to reviewer comments.

%% LyX 2.2.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,times]{extarticle}
\usepackage{mathptmx}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{geometry}
\geometry{verbose,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2.5cm}
\usepackage{array}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[authoryear]{natbib}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\newcommand{\nicefrac}[2]{\ensuremath ^{#1}\!\!/\!_{#2}}
\usepackage { fancybox}
\usepackage[export]{adjustbox}

\usepackage{todonotes}
%\usepackage{afterpage}

%\usepackage[switch]{lineno}

\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

%\usepackage{moreverb}

\usepackage{flushend}

\makeatother

@@

\title{Multi-fluids for Representing Sub-filter-scale Buoyant Convection}

\author{Hilary Weller, William McIntyre, Daniel Shipley and Peter Clark}
\maketitle
\begin{abstract}
Multi-fluid modelling is a promising new method of representing sub
and near grid-scale convection allowing for consistent and accurate
treatment of net mass transport by convection, non-equilibrium and
turbulence. The air is partitioned into convecting and non-convecting
fluids each with their own velocity, temperature and constituents
so that separate equations of motion can be solved for each fluid,
thus representing features that cannot be resolved in a single fluid
at coarse resolution.

@@

@@

Slow progress has been made improving the parameterisation of sub-grid-scale
convection since \citet{AS74,Tied89,GR90,KF90}. Convection is still
one of the weakest aspects of large scale models of the atmosphere
\citep{HPB+14,SAB+13,ipcc41}. Improvements have been made relaxing
the quasi-equilibrium assumption \citep{PR98,GG05,Par14}, including
stochasticity \citep{PC08} and allowing net mass transport by convection
\citep{KB08,MB19}. These developments have enabled limited simulation
in the grey zone, where convection is partially resolved. However
there are still systematic biases implying that improvements should
still be possible.

\citet{TWV+18} proposed multi-fluid modelling of convection in order
to treat non-equilibrium and net mass transport correctly and consistently
with the same numerics and physics controlling thermals, the stable
environment and downdraught dynamics. \citet{Yano14} derived similar
equations for multiple segments and showed how this is a generalisation
of the mass-flux formulation. The \citet{TKP+18} ``extended eddy
diffusivity mass flux'' or extended EDMF scheme also uses similar
equations but \citet{TKP+18} didn't propose a suitable numerical
method for solving the equations. Multi-fluids promises big improvements
in the representation of sub-grid-scale convection and unification
with the representation of turbulence, particularly at the grey zone.
However, there is a long way to go, particularly as multi-fluids requires
a dedicated model rather than being a stand alone parameterisation
for a dynamical core. \citet{TEB19} and \citet{WM19} presented solutions
of the multi-fluid equations assuming two fluids that share the same
pressure (or that pressure differences are predominantly due to drag
between the fluids). \citet{TEB19} showed that these equations (without
drag) are in fact unstable and can be stabilised by diffusion of vertical
velocity whereas \citet{WM19} stabilised the equations with diffusion
and drag between fluids. Neither of these stabilisation techniques
are acceptable and in this paper we show that there are significant
and sustained pressure differences between the fluids (section \ref{sec:results}).

@@

Many of the challenges of representing convection with the multi-fluid
Navier-Stokes equations \citep[eg those described in][]{WM19} carry
over to the Boussinesq equations; convection is buoyancy dominated
and close to incompressible. The resolved test cases presented in
section \ref{sec:results} give very similar results for Navier-Stokes
and Buossinesq equations. Therefore for simplicity we focus on the
multi-fluid Boussinesq equations without background stratification:
\begin{eqnarray}
\frac{\partial\sigma_{i}}{\partial t}+\nabla\cdot\sigma_{i}\mathbf{u}_{i} & = & \sum_{j\ne i}\left\{ \sigma_{j}S_{ji}-\sigma_{i}S_{ij}\right\} \label{eq:sigma}\\
\frac{D_{i}\mathbf{u}_{i}}{Dt}+\nabla P_{i} & = & b_{i}\mathbf{k}+\nu\nabla^{2}\mathbf{u}_{i}+\sum_{j\ne i}\left\{ \frac{\sigma_{j}}{\sigma_{i}}S_{ji}\left(\mathbf{u}_{ji}^{T}-\mathbf{u}_{i}\right)-S_{ij}(\mathbf{u}_{ij}^{T}-\mathbf{u}_{i})-\mathbf{D}_{ij}\right\} \label{eq:mom}\\
\frac{D_{i}b_{i}}{Dt} & = & \alpha\nabla^{2}b_{i}+\sum_{j\ne i}\left\{ \frac{\sigma_{j}}{\sigma_{i}}S_{ji}\left(b_{ji}^{T}-b_{i}\right)-S_{ij}\left(b_{ij}^{T}-b_{i}\right)\right\} \label{eq:b}\\
\sum_{i}\nabla\cdot\sigma_{i}\mathbf{u}_{i} & = & \nabla\cdot\sum_{i}\sigma_{i}\mathbf{u}_{i}=0\label{eq:divFree}\\
\sum_{i}\sigma_{i} & = & 1\label{eq:sumOne}
\end{eqnarray}
with variables defined in table \ref{tab:defns}.

@@

\citet{TEB19} and \citet{WM19} assumed that each co-located fluid
shared the same pressure, leading to unstable equations that \citet{TEB19}
stabilised using diffusion of vertical velocity and \citet{WM19}
stabilised using mass exchanges or drag between the fluids. \citet{TEB19}
achieved realistic results in single column experiments of convective
boundary layer but the amount of diffusion needed would spoil simulations
in the free troposphere. The stabilisation used by \citet{WM19} meant
that the two fluids tended to move as one which defeats the purpose
of using multi-fluid modelling.

The instabilities reported by \citet{WM19} showed growing divergence
in a fluid with vanishing volume fraction. It is mass convergence
rather than velocity convergence that leads to pressure anomalies
that remove the convergence. Therefore if the volume (or mass) fraction
of fluid is vanishingly low then convergence cannot lead to an anomaly
in the total pressure.

@@

\citet{BN86} proposed a parameterisation for pressure differences
in a multi-phase explosion model that has been widely used in mathematics
and engineering for simulating fluidised beds of granular material
\citep[eg.][]{EHM92} and two compressible fluids \citep[eg.][]{SA99}.
In these methods, advection equations are written for the volume fraction,
$\sigma_{i}$, and conservation equations are written for mass fraction,
$\sigma_{i}\rho_{i}$. In our notation, ignoring transfers between
fluids, the \citet{BN86} equations would be:
\begin{eqnarray}
\frac{\partial\sigma_{i}}{\partial t}+\mathbf{V}\cdot\nabla\sigma_{i} & = & \mu\left(P_{i}-P_{j}\right)\label{eq:BNcompaction}\\
\frac{\partial\sigma_{i}\rho_{i}}{\partial t}+\nabla\cdot\left(\sigma_{i}\rho_{i}\mathbf{u}_{i}\right) & = & 0\label{eq:rhoTransport}
\end{eqnarray}
where $\mathbf{V}$ is the velocity of the interface between fluids,
$\mu$ is the compaction viscosity which controls how quickly the
pressure of each fluid relaxes to the mean pressure and index $j$
refers to the other fluid in a two fluid system. In our Boussinesq
system $\rho_{i}$ does not appear in the continuity equation. Different
authors make different assumptions about the interface velocity but
if we assume that $\mathbf{V}=\sum_{i}\sigma_{i}\mathbf{u}_{i}$ and
$\rho_{i}$ is constant then eqns (\ref{eq:divFree}-\ref{eq:rhoTransport})
imply that $\mu p_{i}=-\sigma_{i}^{2}\nabla\cdot\mathbf{u}_{i}$.
If we assume that $\mathbf{V}=\mathbf{u}_{i}$ for each fluid then
we get $\mu p_{i}=-\sigma_{i}\nabla\cdot\mathbf{u}_{i}$.

\citet{TEB19} stabilised the multi-fluid equations by adding a diffusion
term $\nu\partial w^{2}/\partial z^{2}$ to the RHS of the $w$ equation.
If a fluid perturbation pressure is set to $-\nu\partial w/\partial z$
this is equivalent to the diffusion term in one dimension. However
adding a diffusion term will harm three dimensional simulations of
resolved convection in the free troposphere whereas adding a fluid
perturbation pressure will not. Inspired by both \citet{BN86} and
\citet{TEB19} we assume that the pressure in each fluid is given
by:
\begin{equation}
p_{i}=-\gamma\nabla\cdot\mathbf{u}_{i}\label{eq:Pi_div}
\end{equation}
where $\gamma$ is a free coefficient. Dimensional analysis in section
\ref{subsec:dimAnal} will inform the choice of $\gamma$ and sensitivity
to $\gamma$ will be evaluated in section \ref{sec:results}.

@@

We consider dry air without turbulence so use the dynamic entrainment
described by \citet{HC51,AK67,DBF+13}:
\begin{equation}
S_{ij}=\begin{cases}
-\nabla\cdot\mathbf{u}_{i} & \text{if }\nabla\cdot\mathbf{u}_{i}<0\\
0 & \text{otherwise}
\end{cases}
\end{equation}
which is necessary to stop air in one fluid from accumulating when
vertical motion ceases (at the top of a rising plume or at the bottom
of descending air). We will also test a common form of lateral entrainment
\citep{DBF+13}:
\begin{equation}
\varepsilon=\frac{0.2}{r_{c}}\implies S_{01}=w_{1}\frac{0.2}{r_{c}}
\end{equation}
where $\varepsilon$ is the well know fractional entrainment rate
and $r_{c}$ is the cloud or plume radius.

@@

Pressure differences between the fluids can lead to form drag which
is parameterised following \citet{SW69,RC15,WM19} as: 
\begin{equation}
\mathbf{D}_{ij}=\frac{\sigma_{j}C_{D}}{r_{c}}|\mathbf{u}_{i}-\mathbf{u}_{j}|\left(\mathbf{u}_{i}-\mathbf{u}_{j}\right)\label{eq:dragBubble}
\end{equation}
where $C_{D}$ is a drag coefficient, $r_{c}$ is a cloud radius.
Sensitivity to $C_{D}/r_{c}$is explored in section \ref{sec:results}.

@@

The different fluids may not be well mixed so the fluid transferred
may not have the mean properties of the fluid it is leaving \citep[as was assumed by][]{WM19}.
In fact, the most buoyant air should be transferred from fluid 0 to
fluid 1 and vice versa, and the air with least downward momentum should
be transferred from 0 to 1 and vice versa. The properties of the fluids
transferred should depend on the modelling of sub-grid-scale variability
which is beyond the scope of this paper. In section \ref{sec:results}
we conditionally average high resolution solutions based on the sign
of $w$ so that $w<0$ air is in fluid 0 and $w>0$ air in fluid 1.
Therefore the air that is transferred at the interface will have $w=0$.
For two dimensional multi-fluid simulations we assume that the horizontal
velocity transferred is equal to the mean horizontal velocity of the
fluid transferred from. Therefore
\begin{equation}
\mathbf{u}_{ij}^{T}=\begin{pmatrix}u_{i}\\
v_{i}\\
0
\end{pmatrix}.
\end{equation}
For buoyancy, following \citet{TEB19}, we use:
\begin{equation}
b_{ij}^{T}=\theta_{b}b_{i}+(1-\theta_{b})b_{j}
\end{equation}
and present results using $\theta_{b}=\frac{1}{2}$ and $\theta_{b}=1$.
Note that $\theta_{b}\ne1$ implies that the buoyancy of the fluid
transferred depends on the properties of the receiving fluid which
is not logical and can lead to unbounded values of $b_{i}$ in the
fluid that is loosing mass (as demonstrated in section \ref{sec:results}).

@@

The spatial discretisation follows exactly \citet{WM19}, solving
advective form equations using monotonic, finite volume advection
and C-grid, Lorenz staggering. The time-stepping is Crank-Nicolson
with no off-centering and with deferred correction of explicitly solved
variables. It differs from \citet{WM19} because here we solve the
Boussinesq equations and we need stable solutions for equations with
a separate pressure in each fluid. We use two outer iterations per
time step to update explicitly solved variables. For the first outer
iteration, predicted values for time $n+1$ are set to those from
time level $n$ and are given the superscript $\ell$. For the second
iteration, values at $\ell$ are set to those from the end of the
first iteration. The implicit numerical method for applying the transfers
between fluids is specific for two fluids because it involves the
inversion of $2\times2$ matrices.

@@

The first equation to be solved each outer iteration is eqn (\ref{eq:sigma})
for each $\sigma_{i}$, operator split; first applying advection,
then correcting $\sum_{i}\sigma_{i}=1$ and then applying mass transfers:
\begin{eqnarray}
\sigma_{i}^{\prime} & = & \sigma_{i}^{n}-\frac{\Delta t}{2}\nabla\cdot\left\{ \left(\mathbf{u}_{i}^{n}+\mathbf{u}_{i}^{\ell}\right)\left(\sigma_{i}^{n}\right)_{vL}\right\} \\
\sigma_{i}^{\prime\prime} & = & \sigma_{i}^{\prime}\bigg/\sum_{i}\sigma_{i}^{\prime}\\
S_{ij} & = & \begin{cases}
-\nabla\cdot\mathbf{u}_{i}^{\ell} & \text{if }\nabla\cdot\mathbf{u}_{i}^{\ell}<0\\
0 & \text{otherwise}
\end{cases}\\
\sigma_{i}^{n+1} & = & \sigma_{i}^{\prime\prime}+\Delta t\sum_{j\ne i}\sigma_{j}^{\prime\prime}S_{ji}-\sigma_{i}^{\prime\prime}S_{ij}
\end{eqnarray}
where $\Delta t$ is the time step and the superscripts $\prime$
and $\prime\prime$ denote intermediate values. $\sigma_{i}^{n}$
is interpolated from cell centres onto cell faces using monotonic
van-Leer advection (operator $(\sigma_{i}^{n})_{vL}$) as in \citet{WM19}
and values exclusively at time level $n$ for best accuracy and guaranteed
monotonicity. For calculating the divergence, the normal component
of $\mathbf{u}_{i}$ is needed at cell faces which are prognostic
variables of the C-grid. The $S_{ij}$ are limited to ensure $0<\sigma_{i}<1\ \forall i$.

@@

The final stage of the time step iteration is to reconstruct $\mathbf{u}_{i}$
from $u_{i}$ following \citet{WM19}. The stages calculating $P$,
$p_{i}$ and $\mathbf{u}_{i}$ are repeated twice per time step iteration.
The Poisson and Helmholtz equations are solved to a tolerance of $10^{-6}$.

@@

Two test cases are developed to tune parameters and to validate the
use of multi-fluid equations to represent dry, two-dimensional, sub-grid-scale
convection. The first is transient; the standard rising bubble of
\citet{BF02}. We aim to reproduce the vertical heat transport, the
mean velocity of the rising air, the bouyancy of the rising air and
pressure differences between two fluids in a single column two fluid
model. The second test case is steady state with heat transfer at
the ground and uniform radiative cooling. For both test cases the
reference solution is a single fluid solution of the two-dimensional
Boussinesq equations at high horizontal and vertical resolution.

@@

The test case of \citet{BF02} was designed for fully compressible
models but similar solutions are obtained solving Boussinesq equations.
The Boussinesq version has no background stratification in a domain
of height 10\,km and width 20\,km initially at rest. The buoyant
perturbation is given by:
\begin{equation}
b=\frac{1}{15}\cos^{2}\frac{\pi L}{2}\label{eq:thetaPerturb}
\end{equation}
for $L<1$ where $L=\sqrt{\left(\frac{x-x_{c}}{x_{r}}\right)^{2}+\left(\frac{z-z_{c}}{z_{r}}\right)^{2}}$,
$x_{c}=10\ \text{km}$, $z_{c}=2\ \text{km}$ and $x_{r}=z_{r}=2\ \text{km}$.
100\,m grid spacing is used in the $x$ and $z$ directions for the
two dimensional reference simulation and $\Delta z=100\ \text{m}$
is used in the vertical for the single column simulations. All simulations
use a time-step of 2\,s.

@@

\citet{TEB19} and \citet{WM19} assumed that both fluids would share
the same pressure with differences parameterised as drag. The pressure
in figure \ref{fig:bubble} shows that this is a bad assumption since
the large pressure anomalies are in the rising fluid. Drag would always
act to slow down the fluids whereas the negative pressure gradients
in the rising fluid just below the bubble are acting to accelerate
the rising motion.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
%
% This file includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements, including customized APA reference formatting.
%
% You may copy this file and give it your
% article name, and enter your text.
%
%
% Step 1: Set the \documentclass
%
%

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%

\draftfalse

\journalname{Journal of Advances in Modeling Earth Systems (JAMES)}

\usepackage{mathptmx}
\usepackage{array}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}

@@

\title{Multi-fluids for Representing Sub-filter-scale Convection}

\authors
{
    Hilary Weller\affil{1}\thanks{Meteorology, University of Reading, UK}
    ,
    William McIntyre\affil{1}
    ,
    Daniel Shipley\affil{1}
    , and 
    Peter Clark\affil{1}
}

\affiliation{1}{Meteorology, University of Reading, UK}

\correspondingauthor{Hilary Weller}{h.weller@reading.ac.uk}

\begin{keypoints}
\item Multi-fluid modelling enables net mass transport by parameterised convection and non-equlibrium.
\item Multi-fluid modelling means a new dynamical core rather than a parameterisation for an existing core
\item Closure for the pressure difference between fluids and entrainment and detrainment are provided
\end{keypoints}

\begin{abstract}
Multi-fluid modelling is a promising new method of representing sub
and near grid-scale convection allowing for consistent and accurate
treatment of net mass transport by convection, non-equilibrium and
turbulence. The air is partitioned into convecting and non-convecting
fluids each with their own velocity, temperature and constituents
so that separate equations of motion can be solved for each fluid,
thus representing features that cannot be resolved in a single fluid
at coarse resolution.

@@

\section*{Plain Language Summary}
Clouds and buoyant convection are often smaller than the grid size in weather and climate prediction models but they are of central importance. Therefore their effects are parameterised -- their interactions with the larger scales are estimated based on properties of the larger scales. Convection parameterisations are a large source of error in models of the atmosphere despite decades of effort to improve them. An important assumption in most convection parameterisations is that convection moves heat and moisture upwards but not mass. This makes it much simpler to incorporate convection parameterisations into existing models. Multi-fluid modelling will enable mass transport by convection but the approach means a whole new atmospheric model rather than a stand alone routine that can be incorporated into an existing model. This paper provides some closures necessary for multi-fluid modelling and validation of the technique for two dimensional dry convection.

@@

Slow progress has been made improving the parameterisation of sub-grid-scale
convection since \citeA{AS74}, \citeA{Tied89}, \citeA{GR90} and \citeA{KF90}. Convection is still
one of the weakest aspects of large scale models of the atmosphere
\cite{HPB+14,SAB+13,ipcc41}. Improvements have been made relaxing
the quasi-equilibrium assumption \cite{PR98,GG05,Par14}, including
stochasticity \cite{PC08} and allowing net mass transport by convection
\cite{KB08,MB19}. These developments have enabled limited simulation
in the grey zone, where convection is partially resolved. However
there are still systematic biases implying that improvements should
still be possible.

\citeA{TWV+18} proposed multi-fluid modelling of convection in order
to treat non-equilibrium and net mass transport correctly and consistently
with the same numerics and physics controlling thermals, the stable
environment and downdraught dynamics. \citeA{Yano14} derived similar
equations for multiple segments and showed how this is a generalisation
of the mass-flux formulation. The \citeA{TKP+18} ``extended eddy
diffusivity mass flux'' or extended EDMF scheme also uses similar
equations but \citeA{TKP+18} didn't propose a suitable numerical
method for solving the equations. Multi-fluids promises big improvements
in the representation of sub-grid-scale convection and unification
with the representation of turbulence, particularly at the grey zone.
However, there is a long way to go, particularly as multi-fluids requires
a dedicated model rather than being a stand alone parameterisation
for a dynamical core. \citeA{TEB19} and \citeA{WM19} presented solutions
of the multi-fluid equations assuming two fluids that share the same
pressure (or that pressure differences are predominantly due to drag
between the fluids). \citeA{TEB19} showed that these equations (without
drag) are in fact unstable and can be stabilised by diffusion of vertical
velocity whereas \citeA{WM19} stabilised the equations with diffusion
and drag between fluids. Neither of these stabilisation techniques
are acceptable and in this paper we show that there are significant
and sustained pressure differences between the fluids (section \ref{sec:results}).

@@

Many of the challenges of representing convection with the multi-fluid
Navier-Stokes equations \cite<e.g. those described in>{WM19} carry
over to the Boussinesq equations; convection is buoyancy dominated
and close to incompressible. The resolved test cases presented in
section \ref{sec:results} give very similar results for Navier-Stokes
and Buossinesq equations. Therefore for simplicity we focus on the
multi-fluid Boussinesq equations without background stratification:
\begin{eqnarray}
\frac{\partial\sigma_{i}}{\partial t}+\nabla\cdot\sigma_{i}\mathbf{u}_{i} & = & \sum_{j\ne i}\left\{ \sigma_{j}S_{ji}-\sigma_{i}S_{ij}\right\} \label{eq:sigma}\\
\frac{D_{i}\mathbf{u}_{i}}{Dt}+\nabla P_{i} & = & b_{i}\mathbf{k}+\nu\nabla^{2}\mathbf{u}_{i}+\sum_{j\ne i}\left\{ \frac{\sigma_{j}}{\sigma_{i}}S_{ji}\left(\mathbf{u}_{ji}^{T}-\mathbf{u}_{i}\right)-S_{ij}(\mathbf{u}_{ij}^{T}-\mathbf{u}_{i})-\mathbf{D}_{ij}\right\} \label{eq:mom}\\
\frac{D_{i}b_{i}}{Dt} & = & \alpha\nabla^{2}b_{i}+\sum_{j\ne i}\left\{ \frac{\sigma_{j}}{\sigma_{i}}S_{ji}\left(b_{ji}^{T}-b_{i}\right)-S_{ij}\left(b_{ij}^{T}-b_{i}\right)\right\} \label{eq:b}\\
\sum_{i}\nabla\cdot\sigma_{i}\mathbf{u}_{i} & = & \nabla\cdot\sum_{i}\sigma_{i}\mathbf{u}_{i}=0\label{eq:divFree}\\
\sum_{i}\sigma_{i} & = & 1\label{eq:sumOne}
\end{eqnarray}
with variables defined in table \ref{tab:defns}.

@@

\citeA{TEB19} and \citeA{WM19} assumed that each co-located fluid
shared the same pressure, leading to unstable equations that \citeA{TEB19}
stabilised using diffusion of vertical velocity and \citeA{WM19}
stabilised using mass exchanges or drag between the fluids. \citeA{TEB19}
achieved realistic results in single column experiments of convective
boundary layer but the amount of diffusion needed would spoil simulations
in the free troposphere. The stabilisation used by \citeA{WM19} meant
that the two fluids tended to move as one which defeats the purpose
of using multi-fluid modelling.

The instabilities reported by \citeA{WM19} showed growing divergence
in a fluid with vanishing volume fraction. It is mass convergence
rather than velocity convergence that leads to pressure anomalies
that remove the convergence. Therefore if the volume (or mass) fraction
of fluid is vanishingly low then convergence cannot lead to an anomaly
in the total pressure.

@@

\citeA{BN86} proposed a parameterisation for pressure differences
in a multi-phase explosion model that has been widely used in mathematics
and engineering for simulating fluidised beds of granular material
\cite<e.g.>{EHM92} and two compressible fluids \cite<e.g.>{SA99}.
In these methods, advection equations are written for the volume fraction,
$\sigma_{i}$, and conservation equations are written for mass fraction,
$\sigma_{i}\rho_{i}$. In our notation, ignoring transfers between
fluids, the \citeA{BN86} equations would be:
\begin{eqnarray}
\frac{\partial\sigma_{i}}{\partial t}+\mathbf{V}\cdot\nabla\sigma_{i} & = & \mu\left(P_{i}-P_{j}\right)\label{eq:BNcompaction}\\
\frac{\partial\sigma_{i}\rho_{i}}{\partial t}+\nabla\cdot\left(\sigma_{i}\rho_{i}\mathbf{u}_{i}\right) & = & 0\label{eq:rhoTransport}
\end{eqnarray}
where $\mathbf{V}$ is the velocity of the interface between fluids,
$\mu$ is the compaction viscosity which controls how quickly the
pressure of each fluid relaxes to the mean pressure and index $j$
refers to the other fluid in a two fluid system. In our Boussinesq
system $\rho_{i}$ does not appear in the continuity equation. Different
authors make different assumptions about the interface velocity but
if we assume that $\mathbf{V}=\sum_{i}\sigma_{i}\mathbf{u}_{i}$ and
$\rho_{i}$ is constant then eqns (\ref{eq:divFree}-\ref{eq:rhoTransport})
imply that $\mu p_{i}=-\sigma_{i}^{2}\nabla\cdot\mathbf{u}_{i}$.
If we assume that $\mathbf{V}=\mathbf{u}_{i}$ for each fluid then
we get $\mu p_{i}=-\sigma_{i}\nabla\cdot\mathbf{u}_{i}$.

\citeA{TEB19} stabilised the multi-fluid equations by adding a diffusion
term $\nu\partial w^{2}/\partial z^{2}$ to the RHS of the $w$ equation.
If a fluid perturbation pressure is set to $-\nu\partial w/\partial z$
this is equivalent to the diffusion term in one dimension. However
adding a diffusion term will harm three dimensional simulations of
resolved convection in the free troposphere whereas adding a fluid
perturbation pressure will not. Inspired by both \citeA{BN86} and
\citeA{TEB19} we assume that the pressure in each fluid is given
by:
\begin{equation}
p_{i}=-\gamma\nabla\cdot\mathbf{u}_{i}\label{eq:Pi_div}
\end{equation}
where $\gamma$ is a free coefficient. Dimensional analysis in section
\ref{subsec:dimAnal} will inform the choice of $\gamma$ and sensitivity
to $\gamma$ will be evaluated in section \ref{sec:results}.

@@

We consider dry air without turbulence so use the dynamic entrainment
described by \citeA{HC51},\citeA{AK67} and \citeA{DBF+13}:
\begin{equation}
S_{ij}=\begin{cases}
-\nabla\cdot\mathbf{u}_{i} & \text{if }\nabla\cdot\mathbf{u}_{i}<0\\
0 & \text{otherwise}
\end{cases}
\end{equation}
which is necessary to stop air in one fluid from accumulating when
vertical motion ceases (at the top of a rising plume or at the bottom
of descending air). We will also test a common form of lateral entrainment
\cite{DBF+13}:
\begin{equation}
\varepsilon=\frac{0.2}{r_{c}}\implies S_{01}=w_{1}\frac{0.2}{r_{c}}
\end{equation}
where $\varepsilon$ is the well know fractional entrainment rate
and $r_{c}$ is the cloud or plume radius.

@@

Pressure differences between the fluids can lead to form drag which
is parameterised following \citeA{SW69}, \citeA{RC15} and \citeA{WM19} as: 
\begin{equation}
\mathbf{D}_{ij}=\frac{\sigma_{j}C_{D}}{r_{c}}|\mathbf{u}_{i}-\mathbf{u}_{j}|\left(\mathbf{u}_{i}-\mathbf{u}_{j}\right)\label{eq:dragBubble}
\end{equation}
where $C_{D}$ is a drag coefficient, $r_{c}$ is a cloud radius.
Sensitivity to $C_{D}/r_{c}$is explored in section \ref{sec:results}.

@@

The different fluids may not be well mixed so the fluid transferred
may not have the mean properties of the fluid it is leaving \cite<as was assumed by>{WM19}.
In fact, the most buoyant air should be transferred from fluid 0 to
fluid 1 and vice versa, and the air with least downward momentum should
be transferred from 0 to 1 and vice versa. The properties of the fluids
transferred should depend on the modelling of sub-grid-scale variability
which is beyond the scope of this paper. In section \ref{sec:results}
we conditionally average high resolution solutions based on the sign
of $w$ so that $w<0$ air is in fluid 0 and $w>0$ air in fluid 1.
Therefore the air that is transferred at the interface will have $w=0$.
For two dimensional multi-fluid simulations we assume that the horizontal
velocity transferred is equal to the mean horizontal velocity of the
fluid transferred from. Therefore
\begin{equation}
\mathbf{u}_{ij}^{T}=\begin{pmatrix}u_{i}\\
v_{i}\\
0
\end{pmatrix}.
\end{equation}
For buoyancy, following \citeA{TEB19}, we use:
\begin{equation}
b_{ij}^{T}=\theta_{b}b_{i}+(1-\theta_{b})b_{j}
\end{equation}
and present results using $\theta_{b}=\frac{1}{2}$ and $\theta_{b}=1$.
Note that $\theta_{b}\ne1$ implies that the buoyancy of the fluid
transferred depends on the properties of the receiving fluid which
is not logical and can lead to unbounded values of $b_{i}$ in the
fluid that is loosing mass (as demonstrated in section \ref{sec:results}).

@@

The spatial discretisation follows exactly \citeA{WM19}, solving
advective form equations using monotonic, finite volume advection
and C-grid, Lorenz staggering. The time-stepping is Crank-Nicolson
with no off-centering and with deferred correction of explicitly solved
variables. It differs from \citeA{WM19} because here we solve the
Boussinesq equations and we need stable solutions for equations with
a separate pressure in each fluid. We use two outer iterations per
time step to update explicitly solved variables. For the first outer
iteration, predicted values for time $n+1$ are set to those from
time level $n$ and are given the superscript $\ell$. For the second
iteration, values at $\ell$ are set to those from the end of the
first iteration. The implicit numerical method for applying the transfers
between fluids is specific for two fluids because it involves the
inversion of $2\times2$ matrices.

@@

The first equation to be solved each outer iteration is eqn (\ref{eq:sigma})
for each $\sigma_{i}$, operator split; first applying advection,
then correcting $\sum_{i}\sigma_{i}=1$ and then applying mass transfers:
\begin{eqnarray}
\sigma_{i}^{\prime} & = & \sigma_{i}^{n}-\frac{\Delta t}{2}\nabla\cdot\left\{ \left(\mathbf{u}_{i}^{n}+\mathbf{u}_{i}^{\ell}\right)\left(\sigma_{i}^{n}\right)_{vL}\right\} \\
\sigma_{i}^{\prime\prime} & = & \sigma_{i}^{\prime}\bigg/\sum_{i}\sigma_{i}^{\prime}\\
S_{ij} & = & \begin{cases}
-\nabla\cdot\mathbf{u}_{i}^{\ell} & \text{if }\nabla\cdot\mathbf{u}_{i}^{\ell}<0\\
0 & \text{otherwise}
\end{cases}\\
\sigma_{i}^{n+1} & = & \sigma_{i}^{\prime\prime}+\Delta t\sum_{j\ne i}\sigma_{j}^{\prime\prime}S_{ji}-\sigma_{i}^{\prime\prime}S_{ij}
\end{eqnarray}
where $\Delta t$ is the time step and the superscripts $\prime$
and $\prime\prime$ denote intermediate values. $\sigma_{i}^{n}$
is interpolated from cell centres onto cell faces using monotonic
van-Leer advection (operator $(\sigma_{i}^{n})_{vL}$) as in \citeA{WM19}
and values exclusively at time level $n$ for best accuracy and guaranteed
monotonicity. For calculating the divergence, the normal component
of $\mathbf{u}_{i}$ is needed at cell faces which are prognostic
variables of the C-grid. The $S_{ij}$ are limited to ensure $0<\sigma_{i}<1\ \forall i$.

@@

The final stage of the time step iteration is to reconstruct $\mathbf{u}_{i}$
from $u_{i}$ following \citeA{WM19}. The stages calculating $P$,
$p_{i}$ and $\mathbf{u}_{i}$ are repeated twice per time step iteration.
The Poisson and Helmholtz equations are solved to a tolerance of $10^{-6}$.

@@

Two test cases are developed to tune parameters and to validate the
use of multi-fluid equations to represent dry, two-dimensional, sub-grid-scale
convection. The first is transient; the standard rising bubble of
\citeA{BF02}. We aim to reproduce the vertical heat transport, the
mean velocity of the rising air, the bouyancy of the rising air and
pressure differences between two fluids in a single column two fluid
model. The second test case is steady state with heat transfer at
the ground and uniform radiative cooling. For both test cases the
reference solution is a single fluid solution of the two-dimensional
Boussinesq equations at high horizontal and vertical resolution.

@@

The test case of \citeA{BF02} was designed for fully compressible
models but similar solutions are obtained solving Boussinesq equations.
The Boussinesq version has no background stratification in a domain
of height 10\,km and width 20\,km initially at rest. The buoyant
perturbation is given by:
\begin{equation}
b=\frac{1}{15}\cos^{2}\frac{\pi L}{2}\label{eq:thetaPerturb}
\end{equation}
for $L<1$ where $L=\sqrt{\left(\frac{x-x_{c}}{x_{r}}\right)^{2}+\left(\frac{z-z_{c}}{z_{r}}\right)^{2}}$,
$x_{c}=10\ \text{km}$, $z_{c}=2\ \text{km}$ and $x_{r}=z_{r}=2\ \text{km}$.
100\,m grid spacing is used in the $x$ and $z$ directions for the
two dimensional reference simulation and $\Delta z=100\ \text{m}$
is used in the vertical for the single column simulations. All simulations
use a time-step of 2\,s.

@@

\citeA{TEB19} and \citeA{WM19} assumed that both fluids would share
the same pressure with differences parameterised as drag. The pressure
in figure \ref{fig:bubble} shows that this is a bad assumption since
the large pressure anomalies are in the rising fluid. Drag would always
act to slow down the fluids whereas the negative pressure gradients
in the rising fluid just below the bubble are acting to accelerate
the rising motion.
